Sulfur Dioxide on Water's Surface

### 2011-03-21 Background.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

	The results of analyses performed in this study are related to the location of the liquid water surface to relate depth profiles of the orientation distributions. A liquid surface, however, is a dynamic boundary that is neither flat nor stationary. To define a water surface location, several previous studies have used the technique of fitting a lineshape to the density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep. Thus, the averaged density profile fitting will capture average widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes. In this work we approach the problem of defining surface location by calculating a reference location for each timestep. We perform an averaging of the positions of the waters contained in the topmost monolayer. This provides us with a consistent reference point in the simulations to which we relate our analyses. This method provides an intuitive point of reference, but does not increase the computational burden of the analyses.

  The location of the water surface was determined by the averaged positions of a set number of the outter-most waters in the simulated aqueous slabs. The number of waters included in the averaging was decided by taking a few issues into account. First, a visual inspection of the top 30x30 \angs$^2$ cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer. Alternatively, assuming a spherical model of water with a radius of 2.2\angs (the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a similar number of molecules. Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values. Taking the close-packed model as a minimum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating surface location during the simulation analyses.

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. The figure shows a density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work. The water density profile (dashed line) is produced by averaging the density profile of each timestep in the simulation. The water profile was then fit with a $\tanh$ function lineshape to extract the position and width parameters of the water surface. The fitted Gibbs dividing surface has a ``90-10'' width of 8.28 $\AA$. A bulge in the gas-phase (left) side of the water profile is indicative of the drift of the surface waters, or the entire slab over the course of the trajectory. Thus the calculated location and width using the fitting routine are not accurate over long trajectories.

@@

	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region. In order to better understand the chemistry occurring during the \suldiox~adsorption process, the molecular orientation of both the \wat~and \suldiox~must be understood.  The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms. For analysis, a body-fixed frame is defined for \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}. In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell (normal to the plane of the water surface) unless otherwise noted. The orientational analyses presented herein focus on two angles used to define molecular orientation in space.

@@

	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule. $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis. The values of $\phi$ fall in the interval $[-\pi,\pi]$. Given certain values of $\theta$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g. the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular. %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively). $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$. Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$. 
	Figure \ref{fig:water-angles} shows the angle definitions relating the body-fixed axes to a system-fixed reference axis.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/angle-cartoons/molecularframesmall.png}
		\caption{The body-fixed axes are defined with the x-z plane formed by the three atoms, and the z-axis aligned to the molecular bisector. This sets the y-axis normal to the molecular plane, and one of the bonds in the positive x direction.}
		\label{fig:molecular-frame}
	\end{center}
\end{figure}

>>>>

	In this study we create orientational depth-profiles of \suldiox~and \wat. The reference point used for depth calculations is the location of the water surface. A liquid surface, however, is a dynamic boundary that is neither flat nor stationary. To define a water surface location, several previous studies have used the technique of fitting a lineshape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep. Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes. In this work we approach the problem of defining surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer. This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.

  The location of the water surface was determined by the averaged positions of a set number of the outter-most waters in the simulated aqueous slabs. The number of waters included in the averaging was decided by taking a few issues into account. First, a visual inspection of the top cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer. Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a similar number of molecules to the visual inspection. Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values. Taking the close-packed model as a maximum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating surface location during the simulation analyses.

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. The figure shows a density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work. The water density profile (dashed line) is produced by averaging the instantaneous density profile at each timestep in the simulation. The water profile was then fit to a $\tanh$ function (black) to extract the position and width parameters of the water surface. The fitted Gibbs dividing surface has a ``90-10'' width of 8.28 $\AA$. A bulge in the gas-phase (left) side of the water profile is indicative of the drift of the water slab over the course of the trajectory. Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point. All depth locations in our analyses are calculated relative to the surface location of each timestep of the trajectories.

@@

	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region. In order to better understand the chemistry occurring during the \suldiox~adsorption process, the molecular orientation of both the \wat~and \suldiox~must be understood.  The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms. A body-fixed frame is defined for both \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}. In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell normal to the plane of the water surface. The orientational analyses presented herein focus on two angles used to define molecular orientation in space.

@@

	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule. $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis. The values of $\phi$ fall in the interval $[-\pi,\pi]$. Given certain values of $\theta$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g. the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular. %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively). $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$. Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$. 
	However, with $\cos(\theta)$ close to 1 or -1 the symmetry of the plane of the surface results in an isotropic distribution in $\cos(\phi)$. Figure \ref{fig:water-angles} shows the angle definitions relating the body-fixed axes to a system-fixed reference axis.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/angle-cartoons/molecularframesmall.png}
		\caption{Molecular body-fixed axes are defined with the x-z plane formed by the three atoms, and the z-axis aligned to the molecular bisector. The y-axis is normal to the molecular plane, and one of the bonds points in the positive x direction.}
		\label{fig:molecular-frame}
	\end{center}
\end{figure}

### 2011-03-21 Conclusions.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

Our simulations recreated the experimental setup of a recent study by our group,\ref{Ota2011} by introducing a concentrated gas phase of \suldiox~to a water surface. It was found that gaseous \suldiox~quickly adsorbs to the water surface, and continues to bind until a complete surface coverage is reached. Surface waters reorient in the presence of adsorbed \suldiox. The waters at the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present. The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation.

The surface behavior of the \suldiox~molecules was also determined. Experimental surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry. The spectra of the surface \suldiox~vibrational modes are still unreported. However, through the simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption. Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity. This was previously reported by Dang.\ref{Baer2010} At a high \suldiox~concentration, the unreacted \suldiox~moves further out of the water phase as more molecules bind on the surface. The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations. Those \suldiox~molecules at or below the surface strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase. The \suldiox~slightly above the water surface loses the orientational preference within 10 $\AA$ and those further from the water are more isotropically oriented.

@@

This study is one of several needed to characterize \suldiox~adsorption and behavior on aqueous surfaces. The atmospheric chemistry driving many important reactions occurs on aerosol and small particle surfaces. To further our understanding of gaseous adsorption we plan to report further simulation results of temperature and chemical constituent effects on atmospherically relevant water surfaces.

>>>>

Our simulations recreated the experimental setup of a recent study by our group,\cite{Ota2011} by introducing a concentrated gas phase of \suldiox~to a water surface. It was found that gaseous \suldiox~quickly adsorbs to the water surface, and continues to bind until a complete surface coverage is reached. Surface waters reorient in the presence of adsorbed \suldiox. The waters at the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present. The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation.

The surface behavior of the \suldiox~molecules was also determined. Experimental surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry, so the spectra of the surface \suldiox~vibrational modes are still unreported. However, through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption. Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity. This was previously reported by Dang.\cite{Baer2010} At a high \suldiox~concentration, the unreacted \suldiox~moves further out of the water phase as more molecules bind on the surface. The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations. Those \suldiox~molecules at or below the surface strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase. The \suldiox~slightly above the water surface loses the orientational preference within 10 $\AA$ and those further from the water are more isotropically oriented.

@@

This study is one of several aiming to characterize \suldiox~adsorption and behavior on aqueous surfaces. The atmospheric chemistry driving many important reactions occurs on aerosol and small particle surfaces. To gain a better understanding of gaseous adsorption we plan to report further computational simulation and experimental results of temperature and chemical constituent effects on atmospherically relevant water surfaces.

### 2011-03-21 DensityResults.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

One measure of surface activity is the spatial distribution of molecules around the interfacial region. The density distribution of both \wat~and\suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations. The \suldiox~in the neat-water system remained at the top slab surface. While the \suldiox~equilibrated in the saturated system, both in the bulk and gas phases accumulated at both the top and bottom surfaces.

The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the surface, $z$.

@@

Both saturated \suldiox~surfaces exhibit accumulations of \suldiox. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox~that does not adsorb to the surface. The coefficient values for the top surface from eq \ref{eq:gaussian} are $\mu_{top}=2.695$, $A_{top}=7.001$, and $\sigma_{top}=3.203$. The bottom surface values are $\mu_{bottom}=-33.846$, $A_{bottom}=6.618$, and $\sigma_{bottom}=3.151$\angs. The added concentration of \suldiox~creates a layer of molecules adsorbed on top of the water surface. The center of this distribution is further into the gas phase than for the single-\suldiox~molecule surface. The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of \suldiox~surface hydrate complex formation. However, without simulating breakble bonds the chemistry that may pull \suldiox~into the bulk water phase, and the subsequent reaction to form ions is absent. We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox at the interface. However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Dang et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.

>>>>

One measure of surface activity is the spatial distribution of molecules around the interfacial region. The density distribution of both \wat~and \suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations. The \suldiox~in the neat-water system remained at the top slab surface. While the \suldiox~equilibrated in the saturated system, both in the bulk and gas phases accumulated at both the top and bottom surfaces.

The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the top surface, $z$.

@@

Both saturated \suldiox~surfaces exhibit accumulations of \suldiox. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox~that does not adsorb to the surface. The coefficient values for the top surface from eq \ref{eq:gaussian} are $\mu_{top}=2.695$, $A_{top}=7.001$, and $\sigma_{top}=3.203$. The bottom surface values are $\mu_{bottom}=-33.846$, $A_{bottom}=6.618$, and $\sigma_{bottom}=3.151$\angs. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of this distribution is further into the gas phase than for the single-\suldiox~molecule surface. The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of \suldiox~surface hydrate complex formation. However, without simulating breakble bonds the chemistry that may pull \suldiox~into the bulk water phase, and the subsequent reaction to form ions is absent. We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox at the interface. However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Dang et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.

### 2011-03-21 EquilibratedOrientation.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

Molecular orientation analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules in the different locations near and within the water surface. Histograms were created to show the angles $\theta$ and $\phi$ relative to positions from the top surface of the aqueous slab. Both \wat~and \suldiox~orientations are considered for both the neat-\wat~and saturated systems.

@@

The results of the analyses for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water and saturated systems during the equilibrated MD simulations. In both systems an orientational preference is found at the slab surfaces where the water is in contact with the gas phase, or the surface \suldiox~molecules. The histograms are arranged with the plots of $\theta$ on the left and $\phi$ on the right columns, and the results for the neat-water and saturated systems are in the top and bottom rows, respectively. The distance axis was shifted such that the top surface is located at 0\angs, and the bottom surface is approx. 30\angs below. It is also noted that $\theta$ values of the bottom surface are inverted with respect to the top surface and should mirror those results. %The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a peak in the distributions around $\cos(\phi)=1$ at the water surfaces. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.

The histograms show overall similarities for both systems in their shapes and intensities. The presence of a saturated layer of adsorbed \suldiox~molecules alters the water orientation, but not necessarily the location depth of the interface. In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs below the surface location. At 5\angs below the surface the water molecules begin to orient with their bisectors within the plane of the interface (perpendicular to the surface normal, $\cos(\theta)\approx 0$). The corresponding location in the plot of $\phi$ shows that those molecules are also mostly flat to the surface ($\cos(\phi)\approx 1$). Moving further out from the bulk and into the gas phase, the distributions show $\cos(\theta)$ increasing. Waters further out from the bulk have fewer bonding interactions, and orient with their hydrogens more towards the gas phase. The bisector pointing further into the gas phase leads to isotropy in the values of $\phi$.

The change in $\cos(\theta)$ is not as rapid at the neat-water surface as in the saturated system. Furthermore, the noise in the distribution above 5\angs (and below approx -34\angs) in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface. This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the \suldiox~molecules. The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicular to the interface. Recent SFG studies by our group indicate that ***<something goes here>***.

The distribution of $\cos(\phi)$ is more sharply defined (i.e. less isotropic) for the neat-water system than for the saturated one. Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface. The $\phi$ distributions quickly become isotropic above a couple angstroms from the surface as the bisectors orient perpendicularly, and below the surface as the bulk water loses any orientational preference.

@@

Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-\wat~and saturated \suldiox~systems. Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similar to the water orientation distributions plots in figure \ref{fig:water-orientation}). The data set for the neat-\wat~system is much smaller as only a single \suldiox~molecule was simulated in the bulk. The resulting distribution plots are thus noisier than the corresponding saturated plots, but effective comparisons can still be drawn.

The location of \suldiox~within the water surface is slightly different between the two equilibrium MD systems (see figure \ref{fig:density}). The single \suldiox~molecule is generally located just below the 0\angs position of the surface, and the distribution of the concentrated system \suldiox~molecules is centered above the surface location. This indicates that in both systems the \suldiox~is highly surface active, but the locations of the peak \suldiox~densities differ. The single molecule spends most of its time bound within the top water monolayer, and often traversing below that location by only a few\angs. The saturated system's surface molecules sit on top of the monolayer. The angular distribution of the single \suldiox~is concentrated primarily in $\cos(\theta)>0$, such that the bisector points out of the water surface. This same distribution occurs in the saturated system for positions below 0\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface. Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2-angles/so2-angles.png}
		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface. Distributions are shown for $\cos(\theta)$ (left column) and $\cos(\phi)$ (right column) for both the neat-\wat~(top) and high \suldiox~concentration (bottom) systems. For both systems the $\theta$ distributions suggest \suldiox~bound to the water surface with the sulfur pointing towards the water slab, and the oxygens pointing to the gas phase. In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.}
		\label{fig:so2-orientation}
	\end{center}
\end{figure}

>>>>

Molecular orientation analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules in the different locations of the water surface. Histograms were created to show depth-profiles of the angles $\theta$ and $\phi$ in the neat-water and saturated equilibrium systems.

@@

The orientation profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations. In both systems an orientational preference is found at the slab surfaces where the water is furthest towards the gas phase. The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right columns. The top surface is located at 0\angs, and the bottom surface is approx. 30\angs below. It is also noted that $\theta$ values of the bottom surface are inverted with respect to the top surface, mirroring those results.

%The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a peak in the distributions around $\cos(\phi)=1$ at the water surfaces. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.

The histograms show overall similarities for both systems in their shapes and intensities. The presence of a saturated layer of adsorbed \suldiox~molecules alters the water orientation, but not necessarily the orientational depth of the interface. In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs~below the surface location. At 5\angs~below the surface the water molecules begin to orient with their bisectors within the plane of the interface (perpendicular to the surface normal, $\cos(\theta)\approx 0$). The corresponding location in the plot of $\phi$ shows that those molecules are also mostly flat to the surface ($\cos(\phi)\approx 1$). Moving further out from the bulk and into the gas phase, the distributions show $\cos(\theta)$ increasing. Waters further out from the bulk have fewer bonding interactions, and orient with their hydrogens more towards the gas phase. The bisector pointing further into the gas phase leads to isotropy in the values of $\phi$.

The change in $\cos(\theta)$ is not as rapid at the neat-water surface as in the saturated system. Furthermore, the noise in the distribution above 5\angs~(and below approx -34\angs) in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface. This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the adsorbed gas molecules. The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicular to the interface. Our experimental SFG studies also predicted this reorienting behavior due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}

The distribution of $\cos(\phi)$ is more sharply defined (i.e. less isotropic) for the neat-water system than for the saturated one. Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface. The $\phi$ distributions quickly become isotropic above a couple angstroms from the surface as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.

@@

Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems. Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similar to the water orientation distributions plots in figure \ref{fig:water-orientation}). Only a single surface is shown in those plots with \suldiox~data from the second surface averaged with the first. The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk. The resulting distribution plots are thus noisier than the corresponding saturated plots, but effective comparisons can still be drawn.

The angular distribution of the single \suldiox~is concentrated primarily in $\cos(\theta)>0$, such that the bisector points out of the water surface. This same distribution occurs in the saturated system for positions below 0\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface. Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2-angles/so2-angles.png}
		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface. Distributions are shown for $\cos(\theta)$ (left column) and $\cos(\phi)$ (right column) for both the neat-water (top) and saturated (bottom) systems. For both systems the $\theta$ distributions show \suldiox~bound to the water surface with the sulfur pointing towards the water slab, and the oxygens pointing to the gas phase. In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.}
		\label{fig:so2-orientation}
	\end{center}
\end{figure}

### 2011-03-21 Introduction.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

Interactions between sulfur species and aqueous systems are the current topic of an active area of research. Sulfur dioxide remains an important industrial product, and also enters the environment naturally through terrestrial processes. \suldiox~acts as a major component of atmospheric pollution and is a precursor to acid rain formation, and cloud nucleation. Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms. We seek a fuller understanding of the interactions between \suldiox~and water interfaces because of the role \suldiox~takes in atmospheric particle and aerosol chemistry.

Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols. We concluded that an \suldiox~surface complex forms with water when an aqueous surface is exposed to \suldiox~gas. A computational study by Dang et al.\cite{Baer2010} then made a series of predictions of the specific nature of the surface complex through classical and ab initio simulations. That work probed the cause of changes in the surface water OH vibrational spectra, and developed a detailed picture of the nature of the \suldiox~surface complex with water.

Our latest experimental work on \suldiox~behavior further expands our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that atmospherically relevant cold temperatures greatly enhance the binding of gaseous \suldiox~to water interfaces, and facilitates the reaction with water to produce dissolved sulfur species. However, the surface binding of unreacted \suldiox~complexes was found to be a completely reversible process. The same study also conluded that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the reaction of \suldiox, but does not affect the binding of the gaseous molecules. The analysis of the resultant experimental spectra suggested that the surface waters reorient in response to the presence of gaseous \suldiox~complexes, with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface. Because of the geometry of the experiment the \suldiox~modes were not possible to analyze.

In this study we look at the orientational depth profiles of water and \suldiox~during, and after, the adsorption process. Spectroscopic experiments can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations. Thus we hope to complement the experimental results of previous studies. We determine the effect of introducing \suldiox~to a \wat~surface on both the \suldiox~and \wat~molecules using classical molecular dynamics simulations. The results of this study are necessary to understand the behavior of adsorbing gaseous complexes, and we help verify the ascertions of our recent experiments. Along with other ongoing studies, we develop a more complete and detailed picture of gaseous adsorption and complex formation on aqueous surface.

>>>>

Interactions between sulfur species and aqueous systems are the current topic of an active area of research. Sulfur dioxide remains an important industrial product, and also enters the environment naturally through terrestrial processes. \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation. Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms. We seek a fuller understanding of the interactions between \suldiox~and water interfaces because of the central role \suldiox~takes in many atmospheric particle and aerosol reactions.

Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Ota2011} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas. A computational study by Dang et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations. That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.

Our latest experimental work on \suldiox~behavior further expands our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water interfaces at atmospherically relevant cold temperatures is greatly enhanced. The increased binding facilitates the reaction with water to produce dissolved sulfur species. However, the surface binding of unreacted \suldiox~complexes was found to be a completely reversible process. The same study also showed that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the reaction of \suldiox, but do not affect the surface binding of the gas molecules. The resultant experimental spectra suggest that the surface waters reorient in response to the presence of gaseous \suldiox, with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface.

In this study we look at the orientational depth profiles of water and \suldiox~during and after the adsorption process. Spectroscopic experiments can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations. Thus we hope to complement the experimental results of previous studies. We determine the effect of introducing \suldiox~to a \wat~surface, and analyze the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations. The results of this study will help to understand the behavior of adsorbing gaseous complexes, and will verify predictions of our recent experiments. Along with other ongoing experiments and computations, we develop a more complete picture of gaseous adsorption and complex formation on an aqueous surface.
